Static Analysis by Path Focusing

### 2011-04-21 report/report.tex
update of the report

\usepackage{mathcomp,amssymb,verimagstudent}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{mathenv}
\usepackage{tikz}
\usepackage{listings}
\usepackage{placeins}

@@

	\subsubsection{Reachability problem}
	Path focusing technique aims to discover which paths to focus on for the
	computation of the fixpoint iterations. 
	Since the multigraph is not constructed explicitly, the different paths
	starting and finishing on node in $P_R$ are not known. 
	The technique proposes to express the focus path as the solution of a
	satisfiability modulo theory (SMT) problem:

@@

@@

	\subsection {Choice of $P_R$}

@@

>>>>

\usepackage{mathcomp,amssymb,verimagstudent}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{mathenv}
\usepackage{tikz}
\usepackage{listings}
\usepackage{placeins}
\usepackage{algorithm}
\usepackage{algpseudocode}

@@

	\subsubsection{Reachability problem}
	\label{reachability}
	Path focusing technique aims to discover which paths to focus on for the
	computation of the fixpoint iterations. 
	Since the multigraph is not constructed explicitly, the different paths
	starting and finishing on node in $P_R$ are not known. 
	The technique proposes to express the focus path as the solution of a
	satisfiability modulo theory (SMT) problem:

@@

	\subsubsection{Expression of the reachability problem}

	$\rho$ express the semantic of the program. We still need to create the
	formula expressing the reachability problem defined in \ref{reachability}.

	The abstract domain $D$ we use is those of convex polyhedra, so an abstract
	value $X \in D$ is simply a system of linear inequalities between the
	different SSA-variables of the program. Then, it is easy to encode the
	property $x \in X$ into an SMT-formula: one simply write the conjunction of
	these inequalities, with the name of the variables in the vector $x$.

	The formula given to the SMT-solver is:

	$$\rho \wedge b_i^s \wedge \bigwedge_{j\neq i,\ j\in P_R} (\neg b_j^s)
	\wedge x_i \in X_{p_i} \wedge \bigvee_{j/(p_i,p_j)\in E} (b_2^d \wedge \neg
	(x_j \in X_{p_j}))$$

	\begin{itemize}
	\item we search a path starting in $p_i$: $b_i^s$ has to be \emph{true}, and
	all other source points are \emph{false}.
	\item the vector $x_i$ of the numerical variables at point $p_i$ is in our
	abstract value $X_{p_i}$
	\item we search a path arriving in a successor $p_j$ of $p_i$, such as the
	vector of the numerical variables $x_j$ is not yet included in the current
	abstract value $X_{p_j}$ of $p_j$. $b_j^d$ is \emph{true}, and the
	conjunction of the inequalities given by $X_{p_j}$ is \emph{false}.
	\end{itemize}

@@

	Path Focusing algorithm is described in Algorithm \ref{pathfocusingalgo}.
\begin{algorithm}
\caption{Path Focusing}\label{pathfocusingalgo}
\begin{algorithmic}[1] 
\Procedure{pathfocusing}{$P$,$E$,$P_R$}
\ForAll {$p \in P_R$}
	\State Compute $Succ(p)$, the set of the successors of $p$ in the multigraph
\EndFor
\State $\rho \gets$ computeRho$(P,E)$
\State $A \gets \emptyset$
\ForAll {$p \in P_R / I_p \neq \emptyset$}
	\State $A \gets A \cup p$
\EndFor
\While{$A \neq \emptyset$}
	\State Select $p_i \in A$
	\State $A \gets A \setminus \{p_i\}$
	\While{true}
		\State $res \gets SmtSolve\left[\rho \wedge b_i^s \wedge
		\displaystyle\bigwedge_{\substack{j\neq i \\
		j\in P_R}} (\neg b_j^s) \wedge x_i \in X_{p_i} \wedge
		\bigvee_{\substack{j \\ p_j\in Succ(p_i)}} \left(b_2^d \wedge \neg (x'_j \in
		X_{p_j})\right)\right]$
		\If {$res = unsat$}
			\State \textbf{break}
		\EndIf
		\State Compute the focus path $e$ from $p_i$ to $p_j$
		\State $Y \gets \tau_e^\#(X_{p_i})$
		\If {$p_j \in P_W$}
			\State $X_{p_j} \gets X_{p_j} \nabla (X_{p_j} \sqcup Y)$
		\Else
			\State $X_{p_j} \gets X_{p_j} \sqcup Y$
		\EndIf
		\State $A \gets A \cup \{p_j\}$
	\EndWhile
\EndWhile
\State Possibly do some narrowing steps
\State Compute $\{X_{p_i},\ i \notin P_R\}$
\State \textbf{return} $\{X_{p_i},\ i \in P\}$
\EndProcedure
\end{algorithmic}
\end{algorithm}

	\subsection {Precision of the analysis}

@@

	In this subsection, we propose an extension of the path focusing technique
	to compute disjunctive invariants, in order to improve precision of the
	analysis.

	\cite{GulwaniZ10} proposes a technique to compute transitive closure of a
	transition system, i.e compute the invariants for a loop having its semantic
	expressed by the transition system.

	A transition system for a control point $p$ is a DNF formula where each
	disjunct is the semantic of a path from $p$ to itself. It is of the form
	$$\bigvee_{1 \leq i \leq n} \tau_{p,i}$$ where $n$ is the number of
	paths, and $\tau_{p,i}$ is the semantic of the $i$-th path from $p$ to itself.

	The technique aims to compute a disjunctive invariant for this control
	point: 
	$$\displaystyle\bigvee_{1 \leq j \leq m} X_{p,j}$$
	where $X_{p,j}$ is a conjunction of linear inequalities, i.e a convex
	polyhedra.

	The essence of the method is to
	choose an integer $\delta \in \{1,..,m\}$, and a mapping function
	$\sigma: \{1,..,m\} \times \{1,..,n\} \mapsto \{1,..,m\}$.
	For each polyhedra of the disjunctive invariant, and for each path in the
	graph, the image of the polyhedra $X_{p,j}$ by the transition
	$\tau_{p,i}$ is joined with $X_{p,\sigma(j,i)}$.

\begin{algorithm}
\caption{Transitive closure}\label{gulwani}
\begin{algorithmic}[1] 
\Procedure{TransitiveClosure}{$\bigvee_{i=1}^n X_{p,i}$}
\ForAll {$j \in \{1,..,m\} \setminus \{\delta\}$}
	\State $X_{p,j} \gets \perp$
\EndFor
\State $X_{p,\delta} \gets Id$
\Repeat
	\For {$i \in \{1,..,n\}$ and $j \in \{1,..,m\}$}
		\State $X_{p,\sigma(j,i)} \gets X_{p,\sigma(j,i)} \sqcup
		\tau_{p,i}(X_{p,j})$
	\EndFor
\Until {no change in $\bigvee_{j=1}^m X_{p,j}$}
\EndProcedure
\end{algorithmic}
\end{algorithm}

There are heuristics to choose $m,\delta$ and $\sigma$ \cite[Section
5]{GulwaniZ10}.

This algorithm requires to enumerate all the paths and to compute their
semantics, which may result in an exponential blowup. In addition, the
\emph{for} loop at line $7$ iterates on all the $n \times m$ values of $(i,j)$:
some of the join operations may be useless, in the sense that the value of 
$X_{p,\sigma(j,i)}$ may not change. Here, one could use the same technique as
Path Focusing to avoid these drawbacks: one can use SMT-solving to find the
paths $\tau_{p,i}$.

\begin{algorithm}
\caption{Transitive closure with implicit transition system}\label{gulwani}
\begin{algorithmic}[1] 
\Procedure{TransitiveClosureImplicit}{$p$}
\ForAll {$j \in \{1,..,m\} \setminus \{\delta\}$}
	\State $X_{j} \gets \perp$
\EndFor
\State $X_{\delta} \gets Id$
\While {true}

		\State $res \gets SmtSolve\left[\rho \wedge b_k^s \wedge
		\bigvee_{j=1}^m (B_j \wedge x \in X_{j})
		\wedge b_k^d
		\wedge \neg \left(\bigvee_{j=1}^m x' \in X_{j}\right)\right]$

	\If {$res = unsat$}
		\State \textbf{break}
	\EndIf
	\State Compute $\tau_{p,i}$ from $res$ 
	\State $j$ is the number such as $B_j=true$
	\State $X_{p,\sigma(j,i)} \gets X_{p,\sigma(j,i)} \sqcup
	\tau_{p,i}(X_{p,j})$
\EndWhile
\EndProcedure
\end{algorithmic}
\end{algorithm}

A difference is that here, we don't know the number of paths. Then, the mapping
function $\sigma$ has to be defined on $\{1,..,m\} \times \N$.
